---
title: "Language model: quijote"
output: html_notebook
---

```{r}
library(tidyverse)
library(data.table)
library(stringi)
library(compiler)
library(magrittr)
library(glue)
```

```{r}
fns <- list.files(path="data", pattern="*.rds",full.names=T)
fns
```


```{r}
dq1605_chapt_df <- readRDS("data/dq1605_chapt_df.rds")
dq1615_chapt_df <- readRDS("data/dq1615_chapt_df.rds")
```


# Probability n-grams

```{r}
dqboth_chapt_words <- dq1605_chapt_df %>% mutate(book="1605") %>%
  bind_rows(dq1615_chapt_df %>% mutate(book="1615")) %>%
  mutate(book=book%>%as.factor()) %>%
  select(book,chapter_num,content) %>%
  #head(2) %>%
  mutate(content=content %>% str_replace_all("([:punct:])"," \\1 ") %>%
           str_squish %>% str_to_lower(),
         word=content %>% str_split(fixed(" "))) %>%
  select(-content) %>%
  unnest(word)
nrow(dqboth_chapt_words)
n_distinct(dqboth_chapt_words$word)
```

Which are punctuation?

```{r}
dqboth_chapt_words %>%
  filter(str_detect(word,"[:punct:]")) %>%
  count(word,sort=T)
```

Filter bad punctuation

```{r}
dqboth_chapt_words_fix_punct <- dqboth_chapt_words %>%
    filter(!(word %in% c("\"","¿","¡","(",")","&","*"))) %>%
  mutate(word=if_else(word %in% c(";","!"), ".", word))

dqboth_chapt_words_fix_punct %>%
  filter(str_detect(word,"[:punct:]")) %>%
  count(word,sort=T) %>%
  print
```

```{r}
getBookCh <- function(book,ch_num) {
  b <- if(book=="1605") dq1605_chapt_df else dq1615_chapt_df
  b %>%
    filter(chapter_num==ch_num) %>%
    pull(content)
}
findPassage <- function(s,snippet,window=50) s%>%str_extract(sprintf(".{%d}%s.{%d}",window,snippet,window))
```

Investigate non-punctiation data. Note: findPassage is only returning the first occurrence, so the repetition in column "passage"" below is a bug.

```{r}
dqboth_numeric_problems <-
dqboth_chapt_words_fix_punct %>% 
  filter(str_detect(word,"[^[:alpha:][:punct:]]")) %>%
  mutate(passage=pmap_chr(list(book,chapter_num,word),
                          ~getBookCh(..1,..2) %>% findPassage(..3,25)))
dqboth_numeric_problems
```

```{r}
dqboth_numeric_problems$word %>% unique %>% sort
```

```{r}
dqboth_numeric_problems_df <- data_frame(prob=dqboth_numeric_problems$word %>% unique %>% sort,
                                 repl=c("Oh","diez","dieciséis",
                                        "1614","veinte",
                                        "veinticinco","cincuenta")) %T>%
  print
```

Apply str_replace_all for each case above:

```{r}
fixNumericProblemsContencs <- function(s) reduce2(dqboth_numeric_problems_df$prob,
                                                   dqboth_numeric_problems_df$repl,
                                                   ~str_replace_all(..1,..2%s+%"(?= )",..3),
                                                   .init=s)
fixNumericProblemsWord <- function(s) reduce2(dqboth_numeric_problems_df$prob,
                                              dqboth_numeric_problems_df$repl,
                                              ~str_replace(..1,..2%s+%"$",..3), # 1614 not replaced w 16
                                              .init=s)
```

Made it fast by vectorizing

```{r}
dqboth_chapt_words_fix_punct %>%
  mutate(wordNew=fixNumericProblemsWord(word)) %>%
  filter(str_detect(word,"[^[:alpha:][:punct:]]"))
```

Fix nums

```{r}
dqboth_chapt_words_fix_punct_num <- dqboth_chapt_words_fix_punct %>%
  mutate(word=word%>%fixNumericProblemsWord) %T>%
  filter(str_detect(word,"[^[:alpha:][:punct:]]"))
```

### Stopped here to go to dentist

Creates list of triplets

```{r}
dqboth_chapt_triplets <- dqboth_chapt_words_fix_punct_num %>%
  mutate(word=pmap_chr(list(word,lead(word),lead(word,2)),
                   function(a,b,c)str_c(a,b,c,sep="|"))) %>%
  separate(word,into=c("w1","w2","w3"),sep=fixed("\\|")) %>%
  filter(!is.na(w1),w1!="",
         !is.na(w2),w2!="",
         !is.na(w3),w3!="") %>%
  #,!str_detect(w1,"[:punct:]")) %>%
  count(w1,w2,w3,sort=T) %>%
  mutate(triplet_number=row_number()) %>%
  arrange(w1,w2,desc(n))
dqboth_chapt_triplets
```

```{r}
object.size(dqboth_chapt_triplets) %>% format(units="MB")
```

Computes probs that (w1,w2) -> w3, This table can do automatic text generation.

```{r}
dqboth_probs <- dqboth_chapt_triplets %>%
  #head(200) %>%
  rename(count=n) %>%
  group_by(w1,w2) %>%
  summarise(w3=list(w3),
            w3_count=list(count),
            w3_prob=list(count/sum(count))) %>%
  ungroup %>%
  unnest(w3,w3_count,w3_prob) %>%
  as.data.table(key=c("w1","w2")) # for fast lookups
dqboth_probs
```

For consumption by Shiny
```{r}
saveRDS(dqboth_probs,"./data/triplets_quijote.rds")
```

Look at a few samples where w1 and w1 are not punctuation

```{r}
dqboth_probs %>%
  filter(!str_detect(w1,"[[:punct:][:digit:]]"),
         !str_detect(w2,"[[:punct:][:digit:]]"),
         str_length(w1)>2)
```

Use data.table for faster lookups

```{r}
# dqboth_probs <- data.table(dqboth_probs,
#                                              key = c("w1", "w2"))
dqboth_probs[w1=="ya"&w2=="me"]
```

Function which retrieves a probable w3 given w1,w2

```{r}
get_filt <- function(dt,my_w1,my_w2) dt[w1==my_w1&w2==my_w2]
get_w3 <- function(dt,my_w1,my_w2) {
  #print(glue::glue("w1={my_w1},w2={my_w2}"))
  df_filt <- get_filt(dt,my_w1,my_w2) %>%
    as_data_frame %>%
    arrange(w3_prob) %>%
    mutate(w3_prob_cum=cumsum(w3_prob))
  
  r <- runif(1)
  # simulate unfair die: get first prob to be above r
  first_prob <- detect_index(df_filt$w3_prob_cum,~.x>=r)
  df_filt$w3[first_prob]
}
```

# May start here. All needed is a dataframe with the following format: w1,w2,w3,w3_prob

```{r}
dqboth_probs <- readRDS("dqboth_probs.rds")
```


```{r}
get_w3(dqboth_probs,"acompañado","de")
```

Generate N words: (w1,w2) -> w3. (w2,w3) -> w4, ...

```{r}
generate_utterance <- function(dt,w1,w2,n=50) {
  drlib::accumulate_while(.x=c(w1,w2),
                          .f=~c(.x[2],do.call(get_w3,c(list(dt),.x))),
                          .p=~length(.x)==2,
                          #.compare=T,
                          .max=n) %>%
    head_while(~length(.x)==2) %>%
    map_chr(2) %>%
    prepend(w1) %>%
    str_c(collapse=" ") %>%
    str_replace_all("\\s+([:punct:])","\\1")
}
```

```{r}
suppressWarnings(generate_utterance(dqboth_probs,"ya","me",1000))
```


