---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(magrittr)
library(pdftools)

source("../source/language_model.R")
```

```{r}
text <- pdf_text("pantagruel.pdf") %>% str_c(collapse=" ") %>% str_squish
text %>% str_length
```

- The first batch of chapters I-XXXIV is the table of contents, modern french
- The second batch is the same, original french
- The third batch is the modern french text: 71:105
- The fourth is the original french text

```{r}
str_extract_all(text%>%str_replace_all(fixed("'"),"’"),
                regex("(chapitre [ivxl]+)|(prologue de l’auteur)",
                      ignore_case=T)) %>% first
```

```{r}
text_split <- text %>%
  str_replace_all(fixed("’"),"'") %>% # weird big quotes
  str_replace_all(fixed("–"),"-") %>% # weird long bars
  str_split(regex(".(?=(chapitre [ivxl]{1,6})|(prologue de l'auteur))",
                  ignore_case=T)) %>%
  first %>%
  .[72:106]
```

```{r}
text_split %>% length
text_split[c(1,2,35)] %>% str_sub(end=100) # prologue, 1st, last chapts
```

Need to separate text from chapter name. Notice the non-greedy (shortest possible match) with the "?" after ".*"

```{r}
text_split_headings <- text_split %>%
  str_extract(regex("(chapitre [ivxl]{1,6}.*?\\.)|(prologue de l'auteur)",
                    ignore_case=T))
head(text_split_headings)
```

```{r}
text_df <- data_frame(headings=text_split_headings,
                      content=map2_chr(text_split,text_split_headings,
                                       ~str_sub(.x,start=str_length(.y)+1)) %>% str_trim) %>%
  mutate(headings=if_else(str_detect(headings,fixed("Prologue")),
                          headings %>% str_c(" - "),
                          headings)) %>%
  separate(headings,into=c("chapter","title"),sep=fixed(" - ")) %>%
  mutate(chapt_num=row_number()-1) %>%
  select(chapt_num,everything())
```

Needs to fix beginning of each content, which is missing a letter and sometimes has [O]

```{r}
text_df$content %>% str_sub(end=20) %>% str_c(collapse="\n") %>% cat
```

```{r}
insert_chars <- c("T","J","G","Q","J","A","U",
                  "", # Mais
                  "P","U","P","A","A","A","L","U","P",
                  "", # J'ai
                  "E","E","A","P","S","P","Q","A","P",
                  "", # Et comment
                  "A","L","C","A","Q","P","M")
text_df_fixed_test <- data_frame(prefix=text_df$content %>% str_trim() %>% str_sub(end=20),
                                 prefix_sans_O=prefix %>% str_replace("^\\[O\\] ",""),
                                 prefix_new=insert_chars %>% str_c(prefix_sans_O))
```

We must clean up [Note_d+]

```{r}
text_df$content %>% str_extract(regex("...note........",ignore_case = T))
```

```{r}
text_df$content %>% str_extract_all("\\[Note_\\d+?\\]") %>% unlist
```


```{r}
text_df_fixed <- text_df %>% 
  mutate(content=content %>%
           str_trim %>%
           str_replace("^\\[O\\] ",""),
         content=insert_chars %>% str_c(content),
         content=content %>% str_replace_all("\\[Note_\\d+?\\]","")) %>%
  add_column(book="pantagruel",.before = 1) %>%
  mutate(length=content %>% str_length)
glimpse(text_df_fixed)
```





```{r}
text_df_fixed %>%
  ggplot(aes(chapt_num,length)) +
  geom_line() +
  geom_smooth()
```

```{r}
chapt_triplet_probs <- text_df_fixed %>%
  get_triplets_script("../Triplet Language Model/data/rabelais.rds")
```


